\LARGE \bf The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and the Sliding Window Filter

### 2015-03-05 Report/project_report.tex
Added Leutenegger RSS paper and words about MSCKF

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds some preset threshold.

\subsection{MSCKF State Parametrization}
We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.

@@

\subsection{MSCKF State Covariance Propagation}
With reference to the partitions defined in (\ref{eq:covariance}), we compute the predicted camera state covariances and IMU-camera cross-correlations as follows:
\begin{align}
    \hat{\Vec{P}}^-_{CC,k+1} &= \hat{\Vec{P}}_{CC,k} \\
    \hat{\Vec{P}}^-_{IC,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{IC,k}
\end{align}
where $\Delta T$ is the IMU sampling period, and the state transition matrix $\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right)$ is obtained from
\begin{align} \label{eq:statetransition}
	\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) &= \Vec{1}_{12} + \Vec{F} \Delta T.
\end{align}

We obtain the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ according to
\begin{align}
	\hat{\Vec{P}}^-_{II,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{II,k} \boldsymbol{\Phi}^T\left(t_k + \Delta T, t_k \right) \notag \\ &+ \Vec{G} \Vec{Q}_I \Vec{G}^T \Delta T.
\end{align}
To improve numerical stability, we found it useful to enforce the positive semi-definiteness of $\hat{\Vec{P}}^-_{II,k+1}$ by making the replacements
\begin{align*}
    P_{ii} \leftarrow |P_{ii}| && \text{and} && P_{ij}, P_{ji} \leftarrow \dfrac{1}{2}\left(P_{ij} + P_{ji}\right), i\neq j,
\end{align*}
where $P_{ij}$ denotes the $i^\text{th}$ row and $j^\text{th}$ column of $\hat{\Vec{P}}^-_{II,k+1}$.

@@

\begin{figure*}
    \centering
    \subfigure[At least three features are visible in $64\%$ of the interval $(500,1000)$. The number of visible features is frequently exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
        \label{fig:comp_1}
    }
    ~
    \subfigure[At least three features are visible in only $56\%$ of the interval $(1215,1715)$. The number of visible features rarely exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
        \label{fig:comp_2}
    }
    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refer to the number of states in the sliding window.}
\end{figure*}

@@

    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$. The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
    \label{fig:comp_3}
    % \vspace{-0.4cm}
\end{figure}

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $(1215,1715)$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Trans. ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rot. ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF       & Trans. ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
        (20-100)            & Rot. ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF         & Trans. ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
         (25)       & Rot. ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3 GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
    % \vspace{-0.4cm}
\end{table}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/kitti/composite-kitti}
    \caption{Sample frames from the four KITTI traverses shown in Figure \ref{fig:kitti}.}
    \label{fig:composite-kitti}
    % \vspace{-0.4cm}
\end{figure*}

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE) and Average Normalized Estimation Error Squared (ANEES) of IMU integration, MSCKF, and SWF on KITTI traverses.}
    \begin{tabular}{llcccc} 
                    &                   & \multicolumn{4}{c}{KITTI Traverse} \B      \\ \cline{3-6}
                    &                   & \texttt{0001} & \texttt{0036} & \texttt{0051} & \texttt{0095}    \T\B    \\ \hline
        IMU Only    & Trans. ARMSE      & 0.7197    & 0.5131    & 0.7834    & 1.039     \T      \\
                    % & Rot. ARMSE        & 0.0025    & 0.0113    & 0.0034    & 0.0046       \\
                    & ANEES             & 0.1630    & 0.0092    & 0.1170    & 0.6254         \B      \\ \hline
        MSCKF       & Trans. ARMSE      & 0.3492    & 0.4401    & 0.7530    & 0.8170  \T      \\
        % (5-Inf)     & Rot. ARMSE        & 0.0111    & 0.0096    & 0.0091    & 0.0345          \\
        (5-Inf)     & ANEES             & 5.103     & 1.826     & 2.031     & 14.98          \B      \\ \hline
        SWF         & Trans. ARMSE      & 0.3372    & 0.3778    & 0.5832    & 0.7196  \T      \\
        % (10)        & Rot. ARMSE        & 0.0113    & 0.0112    & 0.0111    & 0.0154       \\
        (10)        & ANEES             & 358.3     & 703.2     & 1124      & 3767    \B      \\ \hline
    \end{tabular}
    \label{tab:kitti_results}
    % \vspace{-0.4cm}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
Except where otherwise noted (i.e., state parametrization and integration method), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.

\subsection{MSCKF State Parametrization}
We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.

@@

\subsection{MSCKF State Covariance Propagation}
With reference to the partitions defined in (\ref{eq:covariance}), we compute the predicted camera state covariances and IMU-camera cross-correlations as follows:
\begin{align}
    \hat{\Vec{P}}^-_{CC,k+1} &= \hat{\Vec{P}}_{CC,k} \\
    \hat{\Vec{P}}^-_{IC,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{IC,k}
\end{align}
where $\Delta T$ is the IMU sampling period.

The state transition matrix $\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right)$ and the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ are computed according to \cite{Leutenegger:2013:RSS}:
\begin{align} \label{eq:statetransition}
	\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) &= \Vec{1}_{12} + \Vec{F} \Delta T
\end{align}
\begin{align}
	\hat{\Vec{P}}^-_{II,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{II,k} \boldsymbol{\Phi}^T\left(t_k + \Delta T, t_k \right) \notag \\ &+ \Vec{G} \Vec{Q}_I \Vec{G}^T \Delta T.
\end{align}
To improve numerical stability, we found it useful to enforce the positive semi-definiteness of $\hat{\Vec{P}}^-_{II,k+1}$ by making the replacements
\begin{align*}
    P_{ii} \leftarrow |P_{ii}| && \text{and} && P_{ij}, P_{ji} \leftarrow \dfrac{1}{2}\left(P_{ij} + P_{ji}\right), i\neq j,
\end{align*}
where $P_{ij}$ denotes the $i^\text{th}$ row and $j^\text{th}$ column of $\hat{\Vec{P}}^-_{II,k+1}$.

@@

\begin{figure*}
    \centering
    \subfigure[At least three features are visible in $64\%$ of the interval $(500,1000)$. The number of visible features is frequently exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
        \label{fig:comp_1}
    }
    ~
    \subfigure[At least three features are visible in only $56\%$ of the interval $(1215,1715)$. The number of visible features rarely exceeds ten.] {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
        \label{fig:comp_2}
    }
    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refer to the number of states in the sliding window.}

    \vspace{-0.2cm}
\end{figure*}

@@

    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$. The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
    \label{fig:comp_3}
    \vspace{-0.2cm}
\end{figure}

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $(1215,1715)$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Trans. ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rot. ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF       & Trans. ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
        (20-100)            & Rot. ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF         & Trans. ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
         (25)       & Rot. ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3 GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
    \vspace{-0.2cm}
\end{table}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/kitti/composite-kitti}
    \caption{Sample frames from the four KITTI traverses shown in Figure \ref{fig:kitti}.}
    \label{fig:composite-kitti}
    \vspace{-0.2cm}
\end{figure*}

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE) and Average Normalized Estimation Error Squared (ANEES) of IMU integration, MSCKF, and SWF on KITTI traverses.}
    \begin{tabular}{llcccc} 
                    &                   & \multicolumn{4}{c}{KITTI Traverse} \B      \\ \cline{3-6}
                    &                   & \texttt{0001} & \texttt{0036} & \texttt{0051} & \texttt{0095}    \T\B    \\ \hline
        IMU Only    & Trans. ARMSE      & 0.7197    & 0.5131    & 0.7834    & 1.039     \T      \\
                    % & Rot. ARMSE        & 0.0025    & 0.0113    & 0.0034    & 0.0046       \\
                    & ANEES             & 0.1630    & 0.0092    & 0.1170    & 0.6254         \B      \\ \hline
        MSCKF       & Trans. ARMSE      & 0.3492    & 0.4401    & 0.7530    & 0.8170  \T      \\
        % (5-Inf)     & Rot. ARMSE        & 0.0111    & 0.0096    & 0.0091    & 0.0345          \\
        (5-Inf)     & ANEES             & 5.103     & 1.826     & 2.031     & 14.98          \B      \\ \hline
        SWF         & Trans. ARMSE      & 0.3372    & 0.3778    & 0.5832    & 0.7196  \T      \\
        % (10)        & Rot. ARMSE        & 0.0113    & 0.0112    & 0.0111    & 0.0154       \\
        (10)        & ANEES             & 358.3     & 703.2     & 1124      & 3767    \B      \\ \hline
    \end{tabular}
    \label{tab:kitti_results}
    \vspace{-0.2cm}
\end{table}
